---
title: "Mixed-Effects Models"
subtitle: "Random Intercepts and Random Slopes" 
output:
  html_notebook:
    toc: yes
---
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages_needed <- c("ggplot2", # graphics
                     "dplyr",
                     "arm", # display() etc.
                     "MASS",
                     "ggfortify",
                     "nlme",
                     "lme4",
                     "lmerTest"
                     )
pk_to_install <- packages_needed [!( packages_needed %in% rownames(installed.packages())  )]
if(length(pk_to_install)>0 ){
  install.packages(pk_to_install,repos="http://cran.r-project.org")
}
#lapply(packages_needed, require, character.only = TRUE)
library(ggplot2)
library(dplyr)
library(arm)
library(MASS)
library(ggfortify)
library(nlme)
library(lme4)
library(lmerTest)
```
Part Three: Random intercepts vs. random intercept and slope models

**Sleep Example**

Bates et al. "Fitting Linear Mixed-Effects Models Using lme4"
https://cran.r-project.org/web/packages/lme4/vignettes/lmer.pdf

Average reaction time per day for subjects in a sleep deprivation study (Belenkyet al.2003). On day 0 the subjects had their normal amount of sleep. Starting that night they were restricted to three hours of sleep per night for nine days. 
```{r sleep data}
sleepstudy
```

```{r plot sleep data}
ggplot(data=sleepstudy, aes(Days, Reaction)) + 
  geom_point() +
  stat_smooth(method="lm", se=FALSE) + facet_wrap(.~Subject, nrow = 2) +
  xlab("Days of Sleep Deprivation") +
  ylab("Average Reaction Time (ms)") +
  scale_x_continuous(breaks=seq(0, 9, 3))  # Ticks from 0-9, every 3
```
```{r fit and summarize mixed model}
fm1 <- lmer(Reaction ~ Days + (Days | Subject), sleepstudy)
summary(fm1)
```
We are probably not intrinsically interested in comparing the different individuals that participated in the study.

Continues on Bates p.31